Spontaneous Similarity Judgments of Musical Tones played with various Playing Techniques

### 2017-12-05 paper/lagrangeSpontaneousSimilarity2017.tex
working on intro

@@

The contributions of the paper are as follows: 1) provide perceptual data that account for the perception of musical tones played with various playing techniques, 2), demonstrate for the gathered data, the type of playing technique plays an important role for the organization of the data, and 3) propose a perceptually motivated computational of timbre perception that account well for the gathered human judgments.

>>>>

Musical timbre is an interesting notion to study as it is perceptively defined. Thus, by studying musical timbre, one study many facets of the human auditory system which responds to a diverse -- but somewhat more controled than other stimuli such as speech or environemntal sounds -- set of physical stimuli.

A popular view of the kind of stimuli we consider is that they can be described by a spread of energy that evolve through time and across frequency. By following this trend, most researchers translate the negation of the ASA definition of timbre into a more relative definition : musical timbre is about the distribution of intensity across the time / frequency plane.

A great deal of litterature focuses on the instaneous spectrum, \textit{i.e.} the timbre can be defined without considering evolution of the energy distribution trough time. Other studies demonstrated the importance of the onset in the recognition of some musical instruments. Those studies on a rather short term model where the waveform is observed in a range that do not exceed 100 ms.

Interestingly, modern physiological models of the primary mamalian auditory system  .  Dau's model acount for frequency and rate of modulation. Shamma's model account for frequency, rate and bandwidth to fully describes the physiological evidence gathered by studying the auditory system of the ferret. Though, most of the numerical experiments done using this model did not fully demonstrated the usefullness of this third dimension.

Thus, by taking a signal processing view of the matter, one could consider that a first stage of the auditory system proceeds to a rather fixed set of convolutions across time and frequency, projecting the signal into a large set of descriptive features organized across frequency and rate of modulations.

Thus the second step would be rather opportunistic, in the sense that, from this large set of features, one can pick the ones that are the most relevant for the task at hand.

Considering not only the type of musical instrument but also the playing technique is interesting in that matter, as it invite us to consider a notion of dissimilarity and not the

Following .

@@

The contributions of the paper are as follows: 1) provide perceptual data that account for the perception of musical tones played with various playing techniques, 2), demonstrate that, for the gathered data, the type of playing technique plays an important role for the organization of the data, and 3) propose a perceptually motivated computational model of timbre perception that account well for the human judgments studied in this paper.
